Knowledge \& the web: exercise session 6

### 2015-11-22 exercise_6/ex6.tex
Fixed operators

\section{Rule learning}
One thing we tried was to learn classification rules from the training data.
Rule learning algorithms come with different options for several aspects of the algorithms. In the exercise session we experimented with two of these.\\
First, for choosing the conditions of a classification rule, there are two options: one is to add to the rule the condition which increases the accuracy the most, another is to add to the rule the condition which increases the information gain the most.\\
The accuracy criterion option resulted in the following rule set:
\begin{itemize}

\item if Parch > 5.500 then 0  (1 / 0)
\item if Fare = >10 and Sex = male then 0  (170 / 25)
\item if Sex = female then 1  (63 / 197)
\item else 0  (171 / 61)

@@

\item if Sex = male and Fare = >10 then 0  (170 / 25)
\item if Sex = female then 1  (64 / 197)
\item if Fare = 10-20 and Parch ≤ 0.500 then 0  (66 / 7)
\item else 0  (111 / 54)

@@

\item if Parch > 5.500 then 0  (1 / 0)
\item if Fare = >10 then 0  (189 / 47)
\item if Sex = female then 1  (44 / 175)
\item else 0  (171 / 61)

@@

Apart from \emph{rule learning} we also experimented witch decision trees. First we applied the standard decision tree in RapidMiner, it created a too complicated decision tree, as you can see in figure 1.
\begin{figure}[H]
  \centering
  \includesvg[width = 1.5\textwidth]{decisiontreetitanic_original}
  \caption{Original decision tree with maximal depth of 20}
\end{figure}
Next we limited the maximal depth of the decision tree to 3. This gave us figure 2, which is much more usefull.
\begin{figure}[H]
  \centering
  \includesvg[width = 1\textwidth]{decisiontreetitanic}
  \caption{Decision tree with maximal depth of 3}
\end{figure}

>>>>

\section{Rule learning}
One thing we tried was to learn classification rules from the training data. Rule learning algorithms come with different options for several aspects of the algorithms. In the exercise session we experimented with two of these.
\par First, for choosing the conditions of a classification rule, there are two options: one is to add to the rule the condition which increases the accuracy the most, another is to add to the rule the condition which increases the information gain the most. The accuracy criterion option resulted in the following rule set:
\begin{itemize}

\item if Parch $>$ 5.500 then 0  (1 / 0)
\item if Fare $\geq$ 10 and Sex = male then 0  (170 / 25)
\item if Sex = female then 1  (63 / 197)
\item else 0  (171 / 61)

@@

\item if Sex = male and Fare $\geq$ 10 then 0  (170 / 25)
\item if Sex = female then 1  (64 / 197)
\item if Fare = 10-20 and Parch $\leq$ 0.500 then 0  (66 / 7)
\item else 0  (111 / 54)

@@

\item if Parch $>$ 5.500 then 0  (1 / 0)
\item if Fare $\geq$ 10 then 0  (189 / 47)
\item if Sex = female then 1  (44 / 175)
\item else 0 (171 / 61)

@@

Apart from rule set learning, we also experimented witch decision trees. First we applied the standard decision tree in RapidMiner, it created a too complicated decision tree, as you can see in figure 1.

\begin{figure}[H]
  \centering
  \includegraphics[width = 1.5\textwidth]{decisiontreetitanic_original.pdf}
  \caption{Original decision tree with maximal depth of 20}
\end{figure}
Next we limited the maximal depth of the decision tree to 3. This gave us figure 2, which is much more usefull.
\begin{figure}[H]
  \centering
  \includegraphics[width = 1\textwidth]{decisiontreetitanic.pdf}
  \caption{Decision tree with maximal depth of 3}
\end{figure}
